---
title: "Data_viz"
author: "Brandon"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##https://www.datanovia.com/en/blog/ggplot-examples-best-reference/
##http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Scatterplot
```

## Data Visualization and Exploratory Analysis


```{r}
#install.packages("ggplot2")
#install.packages("pheatmap")

library(ggplot2)
library(pheatmap)
library(dplyr)
library(tidyverse)
library("ggpubr")

```

## Create a random matrix and import the intergrated dataset

```{r pressure, echo=FALSE}
df <- matrix(rnorm(1000),nrow=100)
iris <- iris
```

```{r}
pheatmap(df)
pheatmap(iris[1:4])
```
#plot()

https://r-coder.com/plot-r/

```{r}
plot(iris[,2:3])
```

```{r}
plot(iris)
```
```{r}
hist(iris[,2])
```

```{r}
boxplot(Petal.Length~Species,data=iris)
```

# Label, theme and par

```{r}

plot(iris[,2:3] ,
     main="The Iris Scatter Plot",
     ylab="Sepal Length",
     xlab="Sepal Width",
     cex = 3,
     xlim = c(1, 5),
     ylim = c(1,7)
     )

```

#ggplot
```{r}
# Basic scatter plot
ggplot(mpg, aes(cty, hwy)) +
  geom_point(size = 0.5)

# Jittered points
ggplot(mpg, aes(cty, hwy)) +
  geom_jitter(size = 0.5, width = 0.5)

ggplot(mpg, aes(cty, hwy)) +
  geom_count()
```

#histogram

```{r}
# Basic histogram with mean line
ggplot(iris, aes(Sepal.Length)) +
  geom_histogram(bins = 20, fill = "white", color = "black")  +
  geom_vline(aes(xintercept = mean(Sepal.Length)), linetype = 2)

ggplot(iris, aes(Sepal.Length)) +
  geom_histogram(aes(fill = Species, color = Species), bins = 20, 
                 position = "identity", alpha = 0.5) +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```


#Bar plot
```{r}
df <- mtcars %>%
  rownames_to_column() %>%
  as_data_frame() %>%
  mutate(cyl = as.factor(cyl)) %>%
  select(rowname, wt, mpg, cyl)


# Basic bar plots
ggplot(df, aes(x = rowname, y = mpg)) +
  geom_col() +
  rotate_x_text(angle = 45)

# Reorder row names by mpg values
ggplot(df, aes(x = reorder(rowname, mpg), y = mpg)) +
  geom_col()  +
  rotate_x_text(angle = 45)

# Horizontal bar plots, 
# change fill color by groups and add text labels
ggplot(df, aes(x = reorder(rowname, mpg), y = mpg)) +
  geom_col( aes(fill = cyl)) + 
  geom_text(aes(label = mpg), nudge_y = 2) + 
  coord_flip() +
  scale_fill_viridis_d()



```


#line and point
```{r}
# Data
df3 <- data.frame(supp=rep(c("VC", "OJ"), each=3),
                dose=rep(c("D0.5", "D1", "D2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))

# Line plot
ggplot(df3, aes(x = dose, y = len, group = supp)) +
  geom_line(aes(linetype = supp)) 

# Line plot with [oint]
ggplot(df3, aes(x = dose, y = len, group = supp)) +
  geom_line(aes(linetype = supp)) +
  geom_point(aes(shape = supp))


```

```{r Boxplot}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Basic
ggplot(ToothGrowth, aes(dose, len)) +
  geom_boxplot()

# Box plot + violin plot
ggplot(ToothGrowth, aes(dose, len)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2)


# Plot
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(varwidth=T, fill="plum") + 
    labs(title="Box plot", 
         subtitle="City Mileage grouped by Class of vehicle",
         caption="Source: mpg",
         x="Class of Vehicle",
         y="City Mileage")
```

```{r Time Series}

# Data preparation
df <- economics %>%
  select(date, psavert, uempmed) %>%
  gather(key = "variable", value = "value", -date)

head(df, 3)


# Multiple line plot
ggplot(df, aes(x = date, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  theme_minimal()
```


#coorplot
https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html


```{r}
library(corrplot)
x <- cor(iris[1:4])

corrplot(x, method = "circle")

```


```{r}
library(ggpubr)
# Data preparation
housetasks <- read.delim(
  system.file("demo-data/housetasks.txt", package = "ggpubr"),
  row.names = 1
  )
head(housetasks, 4)


# Visualization
ggballoonplot(housetasks, fill = "value")+
  scale_fill_viridis_c(option = "C")

```


#performance Analysis

```{r}
library(PerformanceAnalytics)

chart.Correlation(
  iris[1:4],
  histogram = TRUE,
  method = c("pearson")
)
```



```{r}
# install.packages("ggdendro")
library(ggplot2)
library(ggdendro)
theme_set(theme_bw())

hc <- hclust(dist(USArrests), "ave")  # hierarchical clustering

# plot
ggdendrogram(hc, rotate = TRUE, size = 2)
```

```{r PCA Plot}
library(FactoMineR)
library(factoextra)

res.pca <- prcomp(iris[, -5],  scale = TRUE)

# Use points only
fviz_pca_ind(res.pca, geom="point")

# Gradient color
fviz_pca_ind(res.pca, col.ind="cos2") +
      scale_color_gradient2(low="white", mid="blue",
      high="red", midpoint=0.6)

p <- fviz_pca_ind(res.pca, label="none", habillage=iris$Species,
             addEllipses=TRUE, ellipse.level=0.95)
print(p)




```


```{r PCA Variable}
# Default plot
fviz_pca_var(res.pca)

# Use points and text
fviz_pca_var(res.pca, geom = c("point", "text"))


fviz_pca_biplot(res.pca)

```


# Exercise

1. This is a subset of a count matrix contain the RNA_Seq results of an experiment pertaining to Parathyrioid,  using the row numbers as your gene i.e. row1 = gene1, and col1 = sample 1
2. Identify which genes have not been expressed in the dataset, 
3. Plot a heatmap of the dataset, showing a hierarchical clustering on the both the rows and columns
4. Can you identify which samples belongs to which groups each samples belongs two? 
5. Plot the same heatmap after log transforming your data in base10
6. Can you identify which samples belongs to which groups each samples belongs two? 
7. Sub-setting the data set to first of the two group as previous selected
8. Calculate the average of each genes within each samples
9. Sub-setting the data set to second of the two group as previous selected
10. Calculate the average of each genes within each samples

```{r}
library(parathyroidSE)
data(parathyroidExonsSE)

cm <- assay(parathyroidExonsSE)[1:1000,1:20]
```




