---
title: "Data_viz"
author: "Brandon"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

##https://www.datanovia.com/en/blog/ggplot-examples-best-reference/
##http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Scatterplot
```

## Data Visualization and Exploratory Analysis


```{r}
#install.packages("ggplot2")
#install.packages("pheatmap")

library('ggplot2')
library('pheatmap')
library('dplyr')
library('tidyverse')
library("ggpubr")

```

## Create a random matrix and import the intergrated dataset

```{r pressure, echo=FALSE}
df <- matrix(rnorm(1000),nrow=10)
iris <- iris
```

```{r}
pheatmap(df)
pheatmap(iris[1:4])
```
#plot()

https://r-coder.com/plot-r/

```{r}
plot(iris[,2:3])
```

```{r}
plot(iris)
```
```{r}
hist(iris[,2])

iris[,2] %>% 
```

```{r}
boxplot(Petal.Length~Species,data=iris)

```

# Label, theme and par

```{r}

plot(iris[,2:3] ,
     main="The Iris Scatter Plot",
     ylab="Sepal Length-test",
     xlab="Sepal Width- testing",
     cex = 7,
     xlim = c(1,6),
     ylim = c(1,10)
     )

```

#ggplot

```{r}
# Basic scatter plot
ggplot(mpg, aes(cty, hwy)) +
  geom_point(size = 0.5)

# Jittered points
ggplot(mpg, aes(cty, hwy)) +
  geom_jitter(size = 0.5, width = 0.5)

ggplot(mpg, aes(cty, hwy)) +
  geom_count()
```

#histogram

```{r}
# Basic histogram with mean line
ggplot(iris, aes(Sepal.Length)) +
  geom_histogram(bins = 8, fill = "white", color = "black")  +
  geom_vline(aes(xintercept = mean(Sepal.Length)), linetype = 1)

ggplot(iris, aes(Sepal.Length)) +
  geom_histogram(aes(fill = Species, color = Species), bins = 20, 
                 position = "identity", alpha = 0.3) +
  scale_fill_viridis_d() +
  scale_color_viridis_d()
```


#Bar plot
```{r}
df <- mtcars %>%
  rownames_to_column() %>%
  as_data_frame() %>%
  mutate(cyl = as.factor(cyl)) %>%
  select(rowname, wt, mpg, cyl)

# Basic bar plots
ggplot(df, aes(x = rowname, y = mpg)) +
  geom_col() +
  rotate_x_text(angle = 45)


# Reorder row names by mpg values
ggplot(df, aes(x = reorder(rowname, mpg), y = mpg)) +
  geom_col()  +
  rotate_x_text(angle = 45)
```


```{r}
# Horizontal bar plots, 
# change fill color by groups and add text labels
ggplot(df, aes(x = reorder(rowname, mpg), y = mpg)) +
  geom_col( aes(fill = cyl)) + 
  geom_text(aes(label = mpg), nudge_y = 3) + 
  coord_flip() +
  scale_fill_viridis_d()

```


#line and point
```{r}
# Data
df3 <- data.frame(supp=rep(c("VC", "OJ"), each=3),
                dose=rep(c("D0.5", "D1", "D2"),2),
                len=c(6.8, 15, 33, 4.2, 10, 29.5))

# Line plot
ggplot(df3, aes(x = dose, y = len, group = supp)) +
  geom_line(aes(linetype = supp)) 

# Line plot with [oint]
ggplot(df3, aes(x = dose, y = len, group = supp)) +
  geom_line(aes(linetype = supp)) +
  geom_point(aes(shape = supp))


```

```{r Boxplot}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)

# Basic
ggplot(ToothGrowth, aes(dose, len)) +
  geom_boxplot()

# Box plot + violin plot
ggplot(ToothGrowth, aes(dose, len)) +
  geom_violin(trim = FALSE) +
  geom_boxplot(width = 0.2)

# Plot
g <- ggplot(mpg, aes(class, cty))
g + geom_boxplot(varwidth=T, fill="plum") + 
    labs(title="Box plot", 
         subtitle="City Mileage grouped by Class of vehicle",
         caption="Source: mpg",
         x="Class of Vehicle",
         y="City Mileage")
```

```{r Time Series}

# Data preparation
df <- economics %>%
  select(date, psavert, uempmed) %>%
  gather(key = "variable", value = "value", -date)

head(df, 3)
str(df)

# Multiple line plot
ggplot(df, aes(x = date, y = value)) + 
  geom_line(aes(color = variable), size = 1)+
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  theme_minimal()
```


#coorplot
https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html


```{r}
library(corrplot)
x <- cor(iris[1:4])

corrplot(x, method = "circle")

```


```{r}
library(ggpubr)
# Data preparation
housetasks <- read.delim(
  system.file("demo-data/housetasks.txt", package = "ggpubr"),
  row.names = 1
  )
head(housetasks, 4)


# Visualization
ggballoonplot(housetasks, fill = "value")+
  scale_fill_viridis_c(option = "C")

```


#performance Analysis

```{r}
library(PerformanceAnalytics)

chart.Correlation(
  iris[1:4],
  histogram = TRUE,
  method = c("pearson")
)
```



```{r}
# install.packages("ggdendro")
library(ggplot2)
library(ggdendro)
theme_set(theme_bw())

hc <- hclust(dist(USArrests), "ave")  # hierarchical clustering

# plot
ggdendrogram(hc, rotate = TRUE, size = 2)
```

```{r PCA Plot}
library(FactoMineR)
library(factoextra)

res.pca <- prcomp(iris[, -5],  scale = TRUE)

# Use points only
fviz_pca_ind(res.pca, geom="point")

# Gradient color
fviz_pca_ind(res.pca, col.ind="cos2") +
      scale_color_gradient2(low="white", mid="blue",
      high="red", midpoint=0.6)

p <- fviz_pca_ind(res.pca, label="none", habillage=iris$Species,
             addEllipses=TRUE, ellipse.level=0.95)
print(p)

```


```{r PCA Variable}
# Default plot
fviz_pca_var(res.pca)

# Use points and text
fviz_pca_var(res.pca, geom = c("point", "text"))


fviz_pca_biplot(res.pca)

```

```{r}
library(plotly)

# plot_ly() tries to create a sensible plot based on the information you 
# give it. If you don't provide a trace type, plot_ly() will infer one.
plot_ly(economics, x = ~pop)
plot_ly(economics, x = ~date, y = ~pop)
```


```{r}
# plot_ly() doesn't require data frame(s), which allows one to take 
# advantage of trace type(s) designed specifically for numeric matrices
plot_ly(z = ~volcano)
```


```{r}
# plotly has a functional interface: every plotly function takes a plotly
# object as it's first input argument and returns a modified plotly object
add_lines(plot_ly(economics, x = ~date, y = ~unemploy/pop))

# To make code more readable, plotly imports the pipe operator from magrittr
economics %>% 
  plot_ly(x = ~date, y = ~unemploy/pop, z=~pce) %>% 
  add_lines()

economics %>% plot_ly(x = ~date, y = ~unemploy/pop) %>% add_paths()
```


```{r}
# Attributes defined via plot_ly() set 'global' attributes that 
# are carried onto subsequent traces, but those may be over-written
plot_ly(economics, x = ~date, color = I("black")) %>%
 add_lines(y = ~uempmed,color= I("yellow")) %>%
 add_lines(y = ~psavert, color = I("red"))
```


```{r}
# Attributes are documented in the figure reference -> https://plotly.com/r/reference
# You might notice plot_ly() has named arguments that aren't in this figure
# reference. These arguments make it easier to map abstract data values to
# visual attributes.
p <- plot_ly(palmerpenguins::penguins, x = ~bill_length_mm, y = ~body_mass_g)
p

```

```{r}
p %>% add_lines(linetype = ~species)

p %>% add_paths(linetype = ~species)
```
```{r}
library(parathyroidSE)
data(parathyroidExonsSE)

cm <- assay(parathyroidExonsSE)[1:5000,1:27]
```

# Exercise

1. This is a subset of a count matrix contain the RNA_Seq results of an experiment pertaining to Parathyrioid,  using the row numbers as your gene i.e. row1 = gene1, and col1 = sample 1

2. Identify which genes have not been expressed in the dataset, 
```{r}

rownames(cm) <- seq(1,nrow(cm)) 
notexpressed <- apply(cm,1,sum)

gene <- notexpressed[notexpressed==0]
gene <-as.data.frame(gene)

```

3. Plot a heatmap of the dataset, showing a hierarchical clustering on the both the rows and columns
```{r}
pheatmap(cm, cluster_rows = TRUE,cluster_cols = TRUE)
```

4. Can you identify which samples belongs to which groups each samples belongs two? 
```{r}
No
```

5. Plot the same heatmap after log transforming your data in base10

```{r}
pheatmap(log10(cm+1))
```


6. What other types of plot would you think it will be interesting to explore? 
 -  Boxplot based on sample?
 -  Animation based on transition between heatmap to heatmap
 -  GO and kegg functional enrichment
 -  Kegg Pathway mapping 

Going Further: 
Using the package "nyclights13" from the previous week, any feature you think you can try to extract and analyze? 
ffmepg





