---
title: "Data Tidying and Transformation"
author: "BrandonYEO"
date: "5/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Installing packages and Load pacakges into environment

```{r cars}
#install.packages("dplyr")
library(dplyr)

#install.packages("nycflights13")
library(nycflights13)

```

loading intergrated dataset and create object 

Try on your side with the import csv and import excel from the top right corner 

```{r}
iris <- iris
cars <- cars
airpas <- AirPassengers
```


view all availible dataset on your Rstudio
```{r}
data() 
```

Convert all data to dataframe

```{r}
iris <- as.data.frame(iris)
cars <- as.data.frame(cars)
airpas <- as.data.frame(airpas)
```


Now let us load the actual data

You might also have noticed the row of three (or four) letter abbreviations under the column names. These describe the type of each variable:

int stands for integers.
dbl stands for doubles, or real numbers.
chr stands for character vectors, or strings.
dttm stands for date-times (a date + a time).

```{r}
flights <- as.data.frame(flights)
dim(flights)
str(flights)
```

# The five key dplyr functions that allow you to solve the vast majority of your data manipulation challenges:

1. Pick observations by their values (filter()).
2. Reorder the rows (arrange()).
3. Pick variables by their names (select()).
4. Create new variables with functions of existing variables (mutate()).
5. Collapse many values down to a single summary (summarise()).

#filter()
```{r}
flight_f <- filter(flights, month == 1)
```


or, Using the pipe functions 
```{r}
flight_fm <-  flights %>% filter(month == 1) 

# is the same as  

flight_fm <- filter(flights, month == 1)
```


filter based on carrier 
```{r}
flight_fc <-  flights %>% filter(carrier == "UA") 

# is the same as  

flight_fc <- filter(flights, carrier == "UA")
```


filter based on carrier 
```{r}
flight_ft <-  flights %>% filter(air_time > 300) 

# is the same as  

flight_ft <- filter(flights, air_time > 300)
```

Using and And OR clause
1. AND &
2. OR |

```{r}
flights_fmOR <- flights %>% filter(month == 11 | month == 12)

flights_fmAND <- flights %>% filter(month <12 & month > 3)

```


using the %in%, which select all elements in the vector afterwards 
```{r}
flights_fmIN <- flights %>% filter(month %in% c(10, 11))

```

flights that were not delayed (on arrival or departure) by more than two hours, you could use either of the following two filters:

- Exclamation mark (!) reverse the TRUE and FALSE

```{r}
flights_NotDelay <- flights %>% filter(!(arr_delay > 120 | dep_delay > 120))

flights_Delay <- flights %>% filter((arr_delay > 120 | dep_delay > 120))
```

Find all flights that

Had an arrival delay of two or more hours

1. Flew to Houston (IAH or HOU)
2, Were operated by United, American, or Delta (c(“UA”,“DL”,“AA”))
3. Departed in summer (July, August, and September)
4. Arrived more than two hours late, but didn't leave late
5. Departed between midnight and 6am (inclusive)


#Arrange() works similarly to filter() except that instead of selecting rows, it changes their order.

```{r}
flight_arrange <- flights %>% arrange( year, month, day)
```

To arrrange the data in desc order
```{r}
flight_arrange_desc <- flights %>% arrange( desc(dep_delay))
```

#Due to the memory limitaton in R, all the code below will ont be pass onto a new object 

Using arrange() to sort all missing values
```{r}
flights %>% 
    arrange(desc(is.na(dep_time)),
           desc(is.na(dep_delay)),
           desc(is.na(arr_time)), 
           desc(is.na(arr_delay)),
           desc(is.na(tailnum)),
           desc(is.na(air_time)))
```

 Find the flights that left earliest.

```{r}
flights %>% 
    arrange(dep_time)
```

# Select columns with select() 
select() allows you to rapidly zoom in on a useful subset using operations based on the names of the variables.

```{r}
x <- flights %>% select(year, month, day)
x
```


To get a more summarised look, use the structure, str() function   
```{r}
str(x)
```



Select columns based on names
```{r}
#one criteria
x <- flights %>% select(starts_with("dep"))
str(x)
```


```{r}
#two criteria with an OR clause

x<- flights %>% select(starts_with("dep") | starts_with("arr"))
str(x)
```

Using one_of(), to select items from a list

```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay","nonsense")

x <- flights %>% select(one_of(vars))

str(x)
```
Using the contains Clause
```{r}
x <- flights %>% select(contains("DEP"))

str(x)
```

# Mutate(). 
mutate() always adds new columns at the end of your dataset so we’ll start by creating a narrower dataset so we can see the new variables.

```{r}
dim(flights)
str(flights)
```
```{r}
x <- flights %>% mutate(new_col = hour*2) 
dim(x)
str(x)
```

Lets use get a closer look
```{r}
y <-  flights %>% 
            mutate(new_col = hour*2) %>% 
            select(hour,new_col)
str(y)

```
```{r}
##################################################
flights_sml <- select(flights, 
                        year:day, 
                        ends_with("delay"), 
                        distance, 
                        air_time
                      )

x <- mutate(flights_sml,
          gain = dep_delay - arr_delay,
          speed = distance / air_time * 60
        )

str(flights_sml)

str(x)
```
```{r}
x <- flights %>%
      select( 
                  year:day, 
                  ends_with("delay"), 
                  distance, 
                  air_time
            )%>%
      mutate(flights_sml,
          gain = dep_delay - arr_delay,
          speed = distance / air_time * 60
        )

str(x)
```

transmute does not preserve the original col

```{r}
x <- flights %>% transmute(
              new_hour = dep_time %/% 100,
              new_minute = dep_time %% 100
            )

str(x)
```


#Summarise(),  It collapses a data frame to a single row:
```{r}
x <- flights %>% summarise(delay = mean(dep_delay, na.rm = TRUE))
x
```

```{r}

x <- flights %>%
      group_by(year, month, day) %>%
      summarise(delay = mean(dep_delay, na.rm = TRUE))
x

dim(x)
str(x)
```

# Workshop Exercise
```{r}
library(parathyroidSE)
data(parathyroidExonsSE)

cm <- assay(parathyroidExonsSE)[1:5000,]

sample <- as.data.frame(colData(parathyroidExonsSE))
```

1. Using the same data from last week, would you able to get a summary of the average gene expression value from each patients? You might need to join the name of the sample to your 

2. from each time frame
3. from each sample 

4, Can you make a new column combining all 3 parameters using the mutate() functions? Do you think you should use mutate()? or should you try to use transmute()?

5. Can you do the same data transformation above now using the new column? You should try the select(contains()) functions from above,   

6. Select sample 1 into a new dataframe and Try to re-arrange the count number from high to low in sample 1 using the arrange command. 

7. can you count the number of genes in sample 1 into 4 categories 1-100, 100-1000, 1000-10000, 10000-anything larger than 10000. You can named them as categories ABCD

8. 







