---
title: "Week 5 Tidying Data"
output: html_notebook
---

# Part One - UBS

```{r}
library(alr4)
library(tidyverse)
library(ggplot2)

```


```{r}
UBSprices <- as.data.frame(UBSprices)
head(UBSprices)
```

#Questions to Answer
#  1. What is the general trend og rice prices in all countries from 2003 to 2009, 
#  use a boxplot to illustrate your findings 

```{r}
tidy_ubs <- UBSprices[,c(3,6)]

tidy_ubs2 <- gather(data = tidy_ubs, key = year, value = price, rice2003, rice2009)

head(tidy_ubs2)

tidy_ubs2$year <- parse_number(tidy_ubs2$year)
tidy_ubs2$year  <- as.factor(tidy_ubs2$year)

ggplot(data=tidy_ubs2,aes(x=year, y=price)) + 
        geom_boxplot()
```

#2. Subset a list of countries that have an increase of rice price from 2003 to 2009? 
#  plot the countries in a bar chart showing the differences in prices
#  subet another list of countries that have an decrease of price from 2003 to 2009

```{r}
tidy_ubs

tidy_ubs$diff <- tidy_ubs$rice2009 - tidy_ubs$rice2003

increase_rice <- tidy_ubs[tidy_ubs$diff>0,]
increase_rice$countries <- as.factor(rownames(increase_rice))

ggplot(data=increase_rice,aes(x=diff,y=countries)) + 
  geom_col()
```

#reordering countries by differences 

```{r}
increase_rice$countries <- fct_reorder(increase_rice$countries, increase_rice$diff) 

ggplot(data=increase_rice,aes(x=diff,y=countries)) + 
  geom_col()

```


##countries with decreased riceprice
```{r}

decrease_rice <- tidy_ubs[tidy_ubs$diff<0,]

decrease_rice$countries <- as.factor(rownames(decrease_rice))

ggplot(data=decrease_rice,aes(x=diff,y=countries)) + 
  geom_col()
```

# reordering countries by differences 

```{r}
decrease_rice$countries <- fct_reorder(decrease_rice$countries, decrease_rice$diff) 

ggplot(data=decrease_rice,aes(x=diff,y=countries)) + 
  geom_col()
```

########################################################################
# Part Two - NYC Flights
########################################################################

```{r}
library(nycflights13)
flights <- flights
str(flights)


#1 Select all flights in November
nov_flights = flights %>% 
                filter(month == 11)

#2 select all flights on 28th of November
nov28 <- flights %>% 
                filter(month == 11, day == 28)

#3 Select from the previous condition, only the columns (flight, origin, distance, air_time)
nov28_4 = nov28 %>% 
              select(flight, origin, distance, air_time)
```


```{r}
#4 Create a new column with the speed of the slight using the question 3
nov28_with_speed = nov28_4 %>% 
              mutate(speed = distance / air_time * 60)

#5. Create a summary of the average speed of all dataset
nov28_with_speed %>%
              summarize(avg_speed = mean(speed,na.rm=TRUE))

#6. Create the same summary of flight from different origins
nov28_with_speed %>% 
              group_by(origin) %>%
              summarize(mean_dist = mean(distance))


```

Calculate destination delay based on destination, and filter the final output

```{r}
delays <- flights %>% 
            group_by(dest) %>% 
            summarise(
              count = n(),
              dist = mean(distance, na.rm = TRUE),
              delay = mean(arr_delay, na.rm = TRUE)
            ) %>% 
               filter(count > 20, dest != "HNL")

library(ggplot2)
ggplot(data = delays, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)

```

# Real World Example
ParathyroidSE is a dataset from bioconduct that contain a RNA-Seq experiment

```{r}
BiocManager::install("parathyroidSE")
library(parathyroidSE)

data(parathyroidGenesSE)
countmatrix <- as.data.frame(parathyroidGenesSE@assays$data$counts)
sample <- as.data.frame(parathyroidGenesSE@colData)
genesname <- as.data.frame(rownames(parathyroidGenesSE))

```

From the data above, extract 
1a. Apply the Columns Name to the count matrix using the colnames() function based on the sample run name

# Some revision from earlier week 
1b. Apply the Row Name to the count matrix using the rownames() function based on the genes names

1c  Extract Sample (RUN) name of DPN treatment on 12h only

1b. Subset the countmatrix to only containing the newly set criteria 

1b. plot a heatmap using pheatmap on the newly extract col information

# Now something more complicated 
2. Can you figure out which sample have the overall lowest count in the sample? plot a box plot of all the sum of all gene to check this out

3. Filter out only the gene that have more than a count of more than 500 in all samples.

4. from the dataset are you able to extract the sample from patient 1, treated with DPB, and plot out the comparison of the two. 

5. Are you able to figure out how many genes have been up regulated and how many have been down regulated? 

6. Can you run a extract the top10 upregulated genes based on fold change? You can calculate fold change of the genes using fc=(expression condition 24h)/(expression condition 48h)

7. Can you do the same for Log2FC? 

8. Are you able to annotate those genes you have isolated earlier? and join them to their Log2FC values

```{r}
## This is an Example to annotate ensemble ID from the dataset, can you modified to fit your list？ 

#BiocManager::install("biomaRt")
library(biomaRt)

mart <- useDataset("hsapiens_gene_ensembl", useMart("ensembl"))
genes <-  genesname

Allattribute <- listAttributes(ensembl)

gene_IDs <- getBM(filters= "ensembl_gene_id", 
                  attributes= c("ensembl_gene_id","hgnc_symbol","description"),
                  values = genes, mart= mart)

Summary <- left_join(genes, gene_IDs, by = c("rownames(parathyroidGenesSE)"="ensembl_gene_id"))

```


8. Are you able to do the same for the other patient? Loops can help you here. 
